% ghcexactprint-Mg.tex
\documentclass[DIV16,twocolumn,10pt]{scrreprt}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage[final]{hcar}

\begin{document}

\begin{hcarentry}[new]{ghc-exactprint}
\report{Matthew Pickering}%05/15
\status{Active, Experimental}
\participants{Alan Zimmerman,Matthew Pickering}% optional
\makeheader

\texttt{ghc-exactprint} aims to be a low-level foundation for
refactoring tools. Unlike most refactoring tools, it works directly with
the GHC API which means that it can understand any legal Haskell source
file.

The program works in two phases. The first phase takes the output from
the parser and converts all absolute source positions into relative
source positions. This means that it is much easier to manipulate the
AST as you do not have to worry about updating irrelevant parts of your
program. The second phase performs the reverse process, it converts
relative source positions back into absolute positions before printing the
source file. The entire library is based around a free monad which keeps track
of which annotations should be where. Each process is then a different
interpretation of this structure.

In theory these two processes should be entirely separate but at the moment they
are not entirely decoupled due to shortcomings we are addressing in GHC 8.0.

In order to verify our foundations, the program has been run on every
source file on Hackage. This testing highlighted a number of bugs which
have been fixed for GHC 7.10.2. Apart from a few outstanding issues with
very rare cases, we can now confidently say that \texttt{ghc-exactprint}
is capable of processing \emph{any} Haskell source file.

Over the last few months Alan Zimmerman has integrated \texttt{ghc-exactprint}
into HaRe\cref{hare} whilst Matthew Pickering participated in Google Summer of
Code to provide integration with \texttt{HLint}.

Both of these proceeded smoothly, and are now working.

\texttt{ghc-exactprint} has also been used for a proof of concept tool to
migrate code forward for AMP and MRP, see link below.

Alan Zimmerman also presented \texttt{ghc-exactprint} at HIW2015, and Matthew
Pickering at SkillsMatter in October. Links to the respective videos are
provided below.

\FurtherReading
\begin{compactitem}
  \item \url{https://github.com/alanz/ghc-exactprint}
  \item \url{https://github.com/hvr/Hs2010To201x}
  \item \url{https://www.youtube.com/watch?v=U5_9mfQAUBo} - HIW2015
  \item
    \url{https://skillsmatter.com/skillscasts/6539-a-new-foundation-for-refactoring-ghc-exactprint}
    - Skills Matter, (free) registration required
\end{compactitem}
\end{hcarentry}

\end{document}